---
title: "dash-smooth and smash-poisson on CTCF ChipExo data"
author: "Kushal K Dey"
date: "4/28/2018"
output: html_document
---

## Introduction

We apply the `dashr-smooth` function on example CTCF Chip-Exo and Chip-seq data, collected by Kevin Luo,
the data are stored under `/project/mstephens/data_smoothing/chipseq/examples` and `/project/mstephens/data_smoothing/chipexo/examples` on Midway cluster. Originally these data have been
downloaded from Pugh et al 2011, Cell, paper. 

We saved some of these example data files under the `data` folder and ran the `dash` and `smash` smoothing 
functions on these data.

```{r}
library(dashr)
library(smashr)
```


## Chip-Exo CTCF

```{r}
chipexo_files <- list.files("../data/chipexo_examples/")
chipexo_files
```

## Processing

```{r}
extract_counts_CTCF <- function(filename){
  bed_counts <- read.table(filename, header = F, stringsAsFactors = F)
  colnames(bed_counts) <- c("chr", "start", "end", "name", "width", "counts")

  counts <- strsplit(bed_counts$counts, split = ",")[[1]]
  counts[counts == "NA"] <- 0
  counts <- as.numeric(counts)

  return(counts.l = list(chr = bed_counts$chr, start = bed_counts$start, end = bed_counts$end, counts = counts))
}
```

## EncodeBroadHistoneHelas - Rep 1

### Forward

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_forward_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(chipexo1$counts)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "CTCF-rep1-forward")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```

### Reverse

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_reverse_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(chipexo1$counts)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "CTCF-rep1-forward")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```

## EncodeBroadHistoneHelas - Rep 2

### Forward

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_forward_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(chipexo1$counts)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "CTCF-rep1-forward")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```

### Reverse

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_reverse_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(chipexo1$counts)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "CTCF-rep1-forward")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```



## EncodeOpenChromChipHelas - Rep 1

### Forward

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_forward_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(chipexo1$counts)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "CTCF-rep1-forward")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```

### Reverse

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_reverse_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(chipexo1$counts)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "CTCF-rep1-forward")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```

## EncodeOpenChromChipHelas - Rep 2

### Forward

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep2_forward_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(chipexo1$counts)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "CTCF-rep1-forward")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```

### Reverse

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep2_reverse_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(chipexo1$counts)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "CTCF-rep1-forward")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```

## Observations

A common observation in all cases is that the dash smoothing is smoother compared to the Poisson SMASH smoothing.
One reason could be that may be taking the logit transformation of the proportional message flow in the Polya tree 
in the SMASH model does not take care of the binomial noise appropriately and causes less smoothing. 

## SessionInfo

```{r}
sessionInfo()
```


