---
title: 'Dash vs SMASH-Poisson comparison : nugget examples'
author: "Kushal K Dey"
date: "4/29/2018"
output: html_document
---

We compare between the smoothing using Dirichlet adaptive shrinkage and the
Poisson SMASH version for simulation examples consisting of nugget effects.

```{r warning = FALSE, message = FALSE}
library(dashr)
library(smashr)
```


## Example 1 - Medium oscillating Poisson nugget

### simulation viz

```{r}
y <- c()
for(num in 1:20){
    y <- c(y, rep(1,10), rep(5, 10))
}
plot(y, type="l")

lambda <- exp(y + rnorm(400, mean=0, sd = 0.5))
plot(lambda, type="l")

```

```{r}
data <- rpois(400, lambda)
data <- data[!is.na(data)]
plot(data, type="l")
length(data)
```

### dash and smash fits

```{r warning = FALSE, message = FALSE}
dash_out <- dashr::dash_smooth(data, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(data)

par(mfrow = c(1,2))
plot(data, col = "gray80")
lines(dash_out$estimate, col = "red", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1),
       cex = 1,
       col=c("gray80","red", "blue"))

plot(data, col = "gray80")
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1),
       cex = 1,
       col=c("gray80", "blue"))

```


## Example 2 - Fast Oscillating Poisson nugget

### simulation viz

```{r}
y <- c()
for(num in 1:100){
    y <- c(y, rep(1,1), rep(5, 1))
}
plot(y, type="l")

lambda <- exp(y + rnorm(400, mean=0, sd = 0.5))
plot(lambda, type="l")

```

```{r}
data <- rpois(400, lambda)
data <- data[!is.na(data)]
plot(data, type="l")
length(data)
```


### dash and smash fits

```{r warning = FALSE, message = FALSE}
dash_out <- dashr::dash_smooth(data, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(data)

par(mfrow = c(1,2))
plot(data, col = "gray80")
lines(dash_out$estimate, col = "red", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1),
       cex = 1,
       col=c("gray80","red", "blue"))

plot(data, col = "gray80")
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1),
       cex = 1,
       col=c("gray80", "blue"))

```



## Example 3 - constant trend Poisson nugget

### simulation viz

```{r}
y <- rep(3,300)
plot(y, type="l")
lambda <- exp(y + rnorm(300, 0, sd = 0.005))
plot(lambda, type="l")

data <- rpois(300, lambda)
data <- data + 1
plot(data, type="l")
```

### dash and smash fits

```{r warning = FALSE, message = FALSE}
dash_out <- dashr::dash_smooth(data, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(data)

par(mfrow = c(1,2))
plot(data, col = "gray80")
lines(dash_out$estimate, col = "red", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1),
       cex = 1,
       col=c("gray80","red", "blue"))

plot(data, col = "gray80")
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1),
       cex = 1,
       col=c("gray80", "blue"))

```

## Example 4 - step trend Poisson nugget

### simulation viz

```{r}
y <- c(rep(3,100), rep(5, 100), rep(6, 100), rep(3, 100), rep(4,100), rep(6, 100))
plot(y, type="l")

lambda <- exp(y + rnorm(600, 0, sd = 0.05))
plot(lambda, type="l")

data <- rpois(600, lambda)
plot(data, type="l")
```


### dash and smash fits

```{r warning = FALSE, message = FALSE}
dash_out <- dashr::dash_smooth(data, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(data)

par(mfrow = c(1,2))
plot(data, col = "gray80")
lines(dash_out$estimate, col = "red", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1),
       cex = 1,
       col=c("gray80","red", "blue"))

plot(data, col = "gray80")
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1),
       cex = 1,
       col=c("gray80", "blue"))


```

