---
title: 'dash : diffeent shrinkage levels CTCF data'
author: "Kushal K Dey"
date: "4/28/2018"
output: html_document
---

## Introduction

Here we try different levels of dash shrinkage on the forward (Watson) and 
the reverse (Crick) strands of the CTCF Chip-exo and Chip-Seq data (rep 1).
The `dash` code allows the user to fix the `Inf_weight` as a control 
parameter which by default is set to 1. We here increase decrease and increase
this parameter and see how that affects the results. 

```{r warning=FALSE, message=FALSE}
library(dashr)
```


## Processing

```{r}
extract_counts_CTCF <- function(filename){
  bed_counts <- read.table(filename, header = F, stringsAsFactors = F)
  colnames(bed_counts) <- c("chr", "start", "end", "name", "width", "counts")

  counts <- strsplit(bed_counts$counts, split = ",")[[1]]
  counts[counts == "NA"] <- 0
  counts <- as.numeric(counts)

  return(counts.l = list(chr = bed_counts$chr, start = bed_counts$start, end = bed_counts$end, counts = counts))
}
```

## CTCF Chip-seq - Rep 1

### Forward

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipseq_examples/example_CTCF_raw_rep1_Forward_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 10), progressbar = FALSE)
out3 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 100), progressbar = FALSE)
```

Plotting the results

```{r}
plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(out$estimate, col = "red", lwd = 2)
lines(out2$estimate, col = "blue", lwd = 2)
lines(out3$estimate, col = "green", lwd = 2)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m-1", "dash-m-10", "dash-m-100"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue", "green"))

```


### Reverse

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipseq_examples/example_CTCF_raw_rep1_Reverse_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 10), progressbar = FALSE)
out3 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 100), progressbar = FALSE)
```

Plotting the results

```{r}
plot(chipexo1$counts, col = "gray80", type = "h", ylab = "reverse strand", xlab = "", main = "CTCF raw rep 1")
lines(out$estimate, col = "red", lwd = 2)
lines(out2$estimate, col = "blue", lwd = 2)
lines(out3$estimate, col = "green", lwd = 2)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m-1", "dash-m-10", "dash-m-100"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue", "green"))

```

## CTCF Chip-Exo - EncodeBroadHistoneHelas - Rep 1

### Forward

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_forward_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 10), progressbar = FALSE)
out3 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 100), progressbar = FALSE)
```

Plotting the results

```{r}
plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF - EncodeBroadHistoneHelas - Rep 1")
lines(out$estimate, col = "red", lwd = 2)
lines(out2$estimate, col = "blue", lwd = 2)
lines(out3$estimate, col = "green", lwd = 2)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m-1", "dash-m-10", "dash-m-100"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue", "green"))

```


### Reverse

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_reverse_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 10), progressbar = FALSE)
out3 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 100), progressbar = FALSE)
```

Plotting the results

```{r}
plot(chipexo1$counts, col = "gray80", type = "h", ylab = "reverse strand", xlab = "", main = "CTCF Chip-Exo - EncodeBroadHistoneHelas - Rep 1")
lines(out$estimate, col = "red", lwd = 2)
lines(out2$estimate, col = "blue", lwd = 2)
lines(out3$estimate, col = "green", lwd = 2)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m-1", "dash-m-10", "dash-m-100"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue", "green"))

```


## CTCF Chip-Exo - EncodeOpenChromChipHelas - Rep 1

### Forward

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_forward_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 10), progressbar = FALSE)
out3 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 100), progressbar = FALSE)
```

Plotting the results

```{r}
plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF - EncodeOpenChromChipHelas - Rep 1")
lines(out$estimate, col = "red", lwd = 2)
lines(out2$estimate, col = "blue", lwd = 2)
lines(out3$estimate, col = "green", lwd = 2)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m-1", "dash-m-10", "dash-m-100"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue", "green"))

```


### Reverse

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_reverse_counts.txt")
out <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 10), progressbar = FALSE)
out3 <- dashr::dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 100), progressbar = FALSE)
```

Plotting the results

```{r}
plot(chipexo1$counts, col = "gray80", type = "h", ylab = "reverse strand", xlab = "", main = "CTCF Chip-Exo - EncodeOpenChromChipHelas - Rep 1")
lines(out$estimate, col = "red", lwd = 2)
lines(out2$estimate, col = "blue", lwd = 2)
lines(out3$estimate, col = "green", lwd = 2)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m-1", "dash-m-10", "dash-m-100"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue", "green"))

```

