---
title: "dash-smooth on consolidated CTCF"
author: "Kushal K Dey"
date: "5/4/2018"
output: html_document
---

## Introduction

We apply the `dashr-smooth` function on example CTCF Chip-Exo and Chip-seq data,
combined across multiple replicates, collected by Kevin Luo,
the data are stored under `/project/mstephens/data_smoothing/chipseq/examples` and `/project/mstephens/data_smoothing/chipexo/examples` on Midway cluster.

```{r warning=FALSE, message = FALSE}
library(dashr)
library(smashr)
```

## Processing

```{r}
extract_counts_CTCF <- function(filename){
  bed_counts <- read.table(filename, header = F, stringsAsFactors = F)
  colnames(bed_counts) <- c("chr", "start", "end", "name", "width", "counts")

  counts <- strsplit(bed_counts$counts, split = ",")[[1]]
  counts[counts == "NA"] <- 0
  counts <- as.numeric(counts)

  return(counts.l = list(chr = bed_counts$chr, start = bed_counts$start, end = bed_counts$end, counts = counts))
}
```

## EncodeBroadHistoneHelas

### Forward strand

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_forward_counts.txt")
chipexo2 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep2_forward_counts.txt")

counts1 <- chipexo1$counts
counts2 <- chipexo2$counts
counts_pool <- counts1 + counts2

out <- dashr::dash_smooth(counts_pool, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(counts_pool)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "EncodeBroadHistoneHelas - forward")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```

### Reverse strand

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_reverse_counts.txt")
chipexo2 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep2_reverse_counts.txt")

counts1 <- chipexo1$counts
counts2 <- chipexo2$counts
counts_pool <- counts1 + counts2

out <- dashr::dash_smooth(counts_pool, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(counts_pool)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "EncodeBroadHistoneHelas-reverse")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```


## EncodeOpenChromChipHelas

### Forward strand

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_forward_counts.txt")
chipexo2 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep2_forward_counts.txt")

counts1 <- chipexo1$counts
counts2 <- chipexo2$counts
counts_pool <- counts1 + counts2

out <- dashr::dash_smooth(counts_pool, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(counts_pool)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "EncodeOpenChromChipHelas - forward")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```



### Reverse strand

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_reverse_counts.txt")
chipexo2 <- extract_counts_CTCF("../data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep2_reverse_counts.txt")

counts1 <- chipexo1$counts
counts2 <- chipexo2$counts
counts_pool <- counts1 + counts2

out <- dashr::dash_smooth(counts_pool, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(counts_pool)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "EncodeOpenChromChipHelas - reverse")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 4)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```

## ChipExo

### Forward strand

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipseq_examples/example_CTCF_raw_rep1_Forward_counts.txt")
chipexo2 <- extract_counts_CTCF("../data/chipseq_examples/example_CTCF_raw_rep2_Forward_counts.txt")
chipexo3 <- extract_counts_CTCF("../data/chipseq_examples/example_CTCF_raw_rep3_Forward_counts.txt")

counts1 <- chipexo1$counts
counts2 <- chipexo2$counts
counts3 <- chipexo3$counts

counts_pool <- counts1 + counts2 + counts3


out <- dashr::dash_smooth(counts_pool, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(counts_pool)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "ChipExo - forward")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```

### Reverse strand

```{r message=FALSE, warning=FALSE}
chipexo1 <- extract_counts_CTCF("../data/chipseq_examples/example_CTCF_raw_rep1_Reverse_counts.txt")
chipexo2 <- extract_counts_CTCF("../data/chipseq_examples/example_CTCF_raw_rep2_Reverse_counts.txt")
chipexo3 <- extract_counts_CTCF("../data/chipseq_examples/example_CTCF_raw_rep3_Reverse_counts.txt")

counts1 <- chipexo1$counts
counts2 <- chipexo2$counts
counts3 <- chipexo3$counts

counts_pool <- counts1 + counts2 + counts3

out <- dashr::dash_smooth(counts_pool, dash_control = list(Inf_weight = 1), progressbar = FALSE)
out2 <- smashr::smash.poiss(counts_pool)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "ChipExo - reverse")
lines(out2, col = "blue", lwd = 2)
lines(out$estimate, col = "red", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(2.5,2.5),
       cex = 0.5,
       col=c("gray80","red", "blue"))

```

