---
title: '*dash-smooth* and *smash-poisson* Comparisons'
author: "Kushal K Dey"
date: "4/28/2018"
output: html_document
---


## Introduction

In this script, we present a few simulation studies to compare the *dash-m* 
(adaptive smoothing using *dash*) and Poisson adaptive smoothing due to 
SMASH (*smashr::smash.poiss*). 

```{r}
library(dashr)
library(smashr)
```


## Example 1 - Poisson with no added noise 

```{r, warning = FALSE, message = FALSE,  fig.align = "left", fig.show="asis", dpi=144, fig.width=6, fig.height=6}
mu <- c(rep(10, 100), rep(20, 100), rep(30, 100), rep(10, 100))
x <- sapply(mu, function(x) rpois(1,x))

out <- dashr::dash_smooth(x, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(x)

plot(x, col = "gray80")
lines(mu, col = "black", lwd = 2)
lines(out$estimate, col = "red", lwd = 2)
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1.5,1.5),
       cex = 1.5,
       col=c("black","red", "blue"))
```

## Example 2 - Poisson with low added noise

```{r, warning = FALSE, message = FALSE, fig.align = "left", fig.show="asis", dpi=144, fig.width=6, fig.height=6}
mu <- c(rep(10, 100), rep(20, 100), rep(30, 100), rep(10, 100))
x <- (mu + rnorm(400, 0, 0.1))
y <- sapply(x, function(x) return(rpois(1,x)))

out <- dashr::dash_smooth(y, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(y)

plot(y, col = "gray80")
lines(mu, col = "black", lwd = 2)
lines(out$estimate, col = "red", lwd = 2)
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1.5,1.5),
       cex = 1.5,
       col=c("black","red", "blue"))
```

## Example 3 -  Poisson GLM model with low noise 

```{r, warning = FALSE, message = FALSE, fig.align = "left", fig.show="asis", dpi=144, fig.width=13, fig.height=12}
mu <- c(rep(1, 100), rep(2, 100), rep(3, 100), rep(1, 100))
x <- (mu + rnorm(400, 0, 0.01))
y <- sapply(x, function(x) return(rpois(1,exp(x))))

out <- dashr::dash_smooth(y, reflect = TRUE)

smash_out <- smashr::smash.poiss(y)

plot(y, col = "gray80")
lines(exp(mu), col = "black", lwd = 4)
lines(out$estimate, col = "red", lwd = 4)
lines(smash_out, col = "blue")
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1.5,1.5),
       cex = 1.5,
       col=c("black","red", "blue"))
```


## Example 4 - A region with consistent high signal 

```{r, warning = FALSE, message = FALSE, fig.align = "left", fig.show="asis", dpi=144, fig.width=6, fig.height=6}
mu <- c(rep(0, 100), rep(3, 30), rep(0, 100), rep(0, 100))
epsilon <- c(rep(0, 100), rnorm(30, 0, 0.01), rep(0, 100), rep(0, 100))
x <- (mu + epsilon)
y <- sapply(x, function(x) return(rpois(1,x)))
ids <- sample(1:length(y), 100, replace = FALSE)
ids <- setdiff(ids, 101:130)
y[ids] <- y[ids] + 1

out <- dashr::dash_smooth(y, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(y)

plot(y, col = "gray80", type = "h")
lines(mu, col = "black", lwd = 2)
lines(out$estimate, col = "red", lwd = 2)
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1.5,1.5),
       cex = 1.5,
       col=c("black","red", "blue"))
```


## Example 4 - A region with inconsistent high signal (interspersed with 0s)

```{r, warning = FALSE, message = FALSE, fig.align = "left", fig.show="asis", dpi=144, fig.width=6, fig.height=6}
mu <- c(rep(0, 100), rep(3, 30), rep(0, 100), rep(0, 100))
epsilon <- rep(0, 330)
ids <- sample(101:130, 10, replace = FALSE)
x = mu + epsilon
y <- sapply(x, function(l) return(rpois(1,l)))
y[ids] = 0

out <- dashr::dash_smooth(y, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(y)

plot(y, col = "gray80", type = "h")
lines(mu, col = "black", lwd = 2)
lines(out$estimate, col = "red", lwd = 2)
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1.5,1.5),
       cex = 1.5,
       col=c("black","red", "blue"))
```


## Example 5 - A region with inconsistent high signal (interspersed with 1 values)

```{r, warning = FALSE, message = FALSE, fig.align = "left", fig.show="asis", dpi=144, fig.width=6, fig.height=6}
mu <- c(rep(0, 100), rep(3, 30), rep(0, 100), rep(0, 100))
epsilon <- rep(0, 330)
ids <- sample(101:130, 10, replace = FALSE)
x = mu + epsilon
y <- sapply(x, function(l) return(rpois(1,l)))
y[ids] = 0
y[101:130] = y[101:130] + 1

out <- dashr::dash_smooth(y, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(y)

plot(y, col = "gray80", type = "h")
lines(mu, col = "black", lwd = 2)
lines(out$estimate, col = "red", lwd = 2)
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1.5,1.5),
       cex = 1.5,
       col=c("black","red", "blue"))
```


## Example 6 - Polynomial curve   + low noise

```{r, warning = FALSE, message = FALSE, fig.align = "left", fig.show="asis", dpi=144, fig.width=6, fig.height=6}
t = seq(-1, 1, length.out = 300)
mu <- 1*t^3 -  2*t + 1
epsilon <- rnorm(300, 0, 0.1)
x = mu + epsilon
y <- sapply(x, function(l) return(rpois(1,exp(l))))


out <- dashr::dash_smooth(y, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(y)

plot(y, col = "gray80", type = "h")
lines(mu, col = "black", lwd = 2)
lines(out$estimate, col = "red", lwd = 2)
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1.5,1.5),
       cex = 1.5,
       col=c("black","red", "blue"))
```


## Example 7 - Polynomial curve  + high noise

```{r, warning = FALSE, message = FALSE, fig.align = "left", fig.show="asis", dpi=144, fig.width=6, fig.height=6}
t = seq(-1, 1, length.out = 300)
mu <- 1*t^3 -  2*t + 1
epsilon <- rnorm(300, 0, 1)
x = mu + epsilon
y <- sapply(x, function(l) return(rpois(1,exp(l))))


out <- dashr::dash_smooth(y, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(y)

plot(y, col = "gray80", type = "h")
lines(mu, col = "black", lwd = 2)
lines(out$estimate, col = "red", lwd = 2)
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1.5,1.5),
       cex = 1.5,
       col=c("black","red", "blue"))
```

## Constant high signal interspersed with 0s

```{r, warning = FALSE, message = FALSE, fig.align = "left", fig.show="asis", dpi=144, fig.width=6, fig.height=6}
y <- rep(3, 300)
ids <- sample(1:300, 100, replace = FALSE)
y[ids] <- 0

out <- dashr::dash_smooth(y, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash_out <- smashr::smash.poiss(y)

plot(y, col = "gray80", type = "h")
lines(out$estimate, col = "red", lwd = 2)
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1.5,1.5),
       cex = 1.5,
       col=c("black","red", "blue"))
```

## Constant high signal interspersed with 1s

```{r, warning = FALSE, message = FALSE, fig.align = "left", fig.show="asis", dpi=144, fig.width=6, fig.height=6}
y <- rep(5, 300)
ids <- sample(1:300, 100, replace = FALSE)
y[ids] <- 1

out <- dashr::dash_smooth(y, dash_control = list(Inf_weight = 1, add_corner = TRUE), progressbar = FALSE)
smash_out <- smashr::smash.poiss(y)

plot(y, col = "gray80", type = "h")
lines(out$estimate, col = "red", lwd = 2)
lines(smash_out, col = "blue", lwd = 2)
legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1.5,1.5),
       cex = 1.5,
       col=c("black","red", "blue"))
```


